---
title: "Heart Disease Analysis"
author: "Rasaq Banjo"
date: '2022-05-17'
output: word_document
df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```
##Import Dataset
```{r}
heart=read.csv("heart disease.csv")
```


##Data Exploration



```{r}
#View first 6 set of the data
head(heart)
```

```{r}
#View last 6 set of the data
tail(heart)
```
```{r}
#Check distribution
table(heart$HeartDisease)
```

```{r}
#Check for missing data
is.na(heart)

```
```{r}
sum(is.na(heart))
```

```{r}
#Dataset summary
summary (heart)
```

```{r}
#Check minimum age
min(heart['Age'])

```

```{r}
#Check Maximum age
max(heart['Age'])

```

```{r}
#Barplot reperesentation of the the dataset

barplot(prop.table(table(heart$HeartDisease)),
        col = rainbow(2),
        ylim = c(0,1),
        main = "Heart Disease Distribution")
```

```{r}
table(heart$HeartDisease,heart$Sex)

barplot(prop.table(table(heart$HeartDisease,heart$Sex)),
        col = rainbow(2),
        ylim = c(0,1),
        main = "Heart Disease Distribution in relation to Sex")
```

##partitioning the data into training and test set

```{r}

library(caret)
trainIndex <- createDataPartition(heart$HeartDisease, p = .67,
                                  list = FALSE)
Train <- heart[ trainIndex,]
Test  <- heart[-trainIndex,]

```


#Using Logistics Regression Model
```{r}
heart=read.csv("heart disease.csv")
xtabs(~HeartDisease +Sex+ ChestPainType, data = Train) # table of categorical outcome

```

```{r}
library(ggplot2)
plot=ggplot(heart, aes(x=Cholesterol , y= RestingBP, color=HeartDisease)) + geom_line()

plot

```

```{r}
ggplot(heart, aes(x=HeartDisease, fill=Sex))+geom_bar()
```


#Logistic model
```{r}
logit <- glm(HeartDisease ~. , data=Train, family = "binomial")

summary(logit)

```


```{r}
confint(logit) # CIs using profiled log-likelihood
```

#Wald Test
```{r}

library(aod)
wald.test(b = coef(logit), Sigma = vcov(logit), Terms = 4:6)


```

```{r}
exp(cbind(OR = coef(logit), confint(logit))) # OR and 95% CI
```

##Likelihood Ratio Test

```{r}
logLik(logit) # see the model's log-likelihood
```

```{r}

with(logit, null.deviance - deviance) # find the difference in deviance
```

```{r}

with(logit, df.null - df.residual) # The df for the difference between the two models = the number of predictor variables
```

```{r}

with(logit, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE)) # obtain p-value
```


##Predicting probability
```{r}
n.n1 <- predict(logit, Test, type = "response")

n.n2 <- with(heart, data.frame(HeartDisease = rep(seq(from = 200, to = 800, length.out = 100),4),Cholesterol = mean(Cholesterol)
                              ,RestingBP=rep(seq(from = 200, to = 800, length.out = 100),4)))

n.n3 <- cbind(heart, predict(logit, newdata =heart, type = "link", se = TRUE)) # including standard error for CI later
n.n3 <- within(n.n3, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})
View(n.n3)
hist(n.n3$PredictedProb, main="Histogram of predicted probabilities", col="red")
```
##Obtaining the test data
```{r}

library(dplyr)
Test <- Test  %>% mutate(model_pred = 1*(n.n1 > .53) + 0,
                                 HeartDisease = 1*(HeartDisease == "Yes") + 0)
```

#Plotting the predicted probabilities 

```{r}
library(ggplot2)
ggplot(n.n3, aes(x = HeartDisease, y = PredictedProb)) + 
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = ChestPainType
  ), alpha = 0.2) + geom_line(aes(colour= Sex), size = 1)
```
##Determining the model accuracy for logistic regression
```{r}
Test <- Test %>% mutate(accurate = 1*(model_pred == HeartDisease))
sum(Test$accurate)/nrow(Test)

```

##Using Decision Tree
```{r}
library(rpart)
Train$HeartDisease <- factor(Train$HeartDisease)

Tree <- rpart(HeartDisease ~. -HeartDisease, data =Train)

plot(Tree, main="Decision Tree of Heart Disease")

text(Tree, pretty = 2)


```


#Showing the Segments of the Anomalies 

```{r}
plot(y=heart$Cholesterol, x=heart$MaxHR, 
     pch = 16, col = "blue", ylab = "Cholesterol", xlab = "Max HR")
abline(v =136.5, col = "red", lwd = 2) #a vertical line

```


##Making prediction

```{r}
library(rpart)
predict_unseen <-predict(Tree, Train, type = 'class')

```

##Testing the patients that did have heart disease and those that didnt

```{r}
table_heart <- table(Train$HeartDisease, predict_unseen)
table_heart

```
##estimating the confusion matrix accuracy metrics

```{r}
accuracy_Test <- sum(diag(table_heart)) / sum(table_heart)
print(paste('Accuracy for test', accuracy_Test))

```















